### 2014-05-15 latex/paper1.tex
i think i'm done

@@

In this model, charged particles orbit a central nucleus in circles.
When a magnetic field is applied, the orbits are perturbed, so that the
frequency of light emitted or absorbed by the atom changes slightly. The
result is that one spectral line splits into three, and therefore looks
as if the line widened. Using Maxwell's electromagnetic theory and
assuming this model of the atom, Lorentz predicted the polarization of
light for each of the three split lines, and his prediction was found to
be in exact agreement with Zeeman's experiment. This was done around the
time, if not before the discovery of the electron by J.\ J.\ Thomson,
showing that there was an intense interest and rush to understand the
structure of matter. Furthermore, this model of the atom along with the
experiment allowed Zeeman to calculate the charge-to-mass ratio of the
electron, which was in agreement with the value obtained by J.\ J.\
Thomson.

@@

>>>>

At the turn of the 20th century, physics had progressed to the point
where it was now possible to investigate the structure of matter and
light. Electromagnetic theory had been developed so that physicists
could investigate the charged particles making up the atom. Statistical
mechanics, originally developed to explain heat and gases, would help
Planck and Einstein realize the corpuscular nature of light. Physics was
ready for a major transformation. In the past, physics had been
primarily concerned with the behavior and motion of things, such as the
motion of planets, the behavior of light, the behavior of electric and
magnetic fields, or the behavior of heat and gases. In this paper, I
argue that at the turn of the 20th century, physics shifted from being
mainly concerned with the behavior of matter and light to being
interested in the actual structure of matter and light. Physicists
became interested in what made up the the things they were studying,
what was inside the atom, what was light made of, and physicists finally
had the tools they needed to find answers to these questions. I will
use the investigations into radioactivity, Thomson's discovery of the
electron, Zeeman's experiments, and Planck and Einstein's ideas about
light to illustrate the heightened interest in the structure of matter
of light during this time period.

@@

In this model, charged particles orbit a central nucleus in circles.
When a magnetic field is applied, the orbits are perturbed, so that the
frequency of light emitted or absorbed by the atom changes slightly. The
result is that one spectral line splits into three, and therefore looks
as if the line widened. Using Maxwell's electromagnetic theory and
assuming this model of the atom, Lorentz predicted the polarization of
light for each of the three split lines, and his prediction was found to
be in exact agreement with Zeeman's experiment. This was done around the
time, if not before the discovery of the electron by J.\ J.\ Thomson,
showing that there was an intense interest and rush to understand the
structure of matter. Furthermore, this model of the atom along with the
experiment allowed Zeeman to calculate the charge-to-mass ratio of the
electron, which was in agreement with the value obtained more directly
by J.\ J.\ Thomson.

@@

These experiments and investigations during the turn of the century
propelled physics into a new era. The intense interest in the atom and
the nature of light created during this time fueled the quantum
revolution taking place twenty years later. Throughout the 20th century,
a large part of physics continued to investigate the structure of matter
and light. Many new particles and families of particles were discovered,
such as neutrinos, quarks, force carriers, and very recently the Higgs
boson. An important breakthrough during the turn of the 20th century
that I did not mention earlier in this paper is Einstein's theory of
relativity. It was one of the few accomplishments during this time that
was not concerned with the structure of matter or light, but was
concerned with the nature of space and time. This theory later lead to
Einstein's theory of gravitation. Interestingly, while all aspects
of physics that contributed to understanding the structure of matter and
light at the time have now been reformulated as quantum theories, such
as electromagnetic theory or statistical mechanics, Einstein's theory of
gravity, which did not come from investigating the structure of matter
or light, to this day has no satisfactory quantum counterpart.
